---
title: "DDP-Wk3-PGA1"
author: "Gerrit Versteeg"
date: "24 July 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Setup environment
First let's setup the appropriate libraries:
```{r init, echo=TRUE, warning=FALSE, error=FALSE}
library("ggplot2", warn.conflicts=FALSE)  ## load ggplot2 silently
library("plotly", warn.conflicts=FALSE)  ## load plotly silently
library("tidyr", warn.conflicts=FALSE)     ## load plyr silently
library("dplyr", warn.conflicts=FALSE)    ## load dplyr silently
```

## The data to be plotted
As an illustration of the use of plotly in the context of the peer graded assignment for week 3 of the DDP course, we will plot the levels of air polution. In The Hague (the Metherlands) these levels are measured by four measuring stations exploited by an organization called RIVM. The data for this research is retreived from <https://www.luchtmeetnet.nl/download> by selecting one of the four RIVM stations, namely the one at the "Rebecquestraat".

 - Export_RS.csv  (468 kB) - Rebecquestraat

The file is in European csv-formaat (separator=";" and decimal=","), so we need to use read.csv2. The file is placed in a tibble (DF_RS). 

## Load the data to be plotted
```{r loading, echo=TRUE, comment=""}
DF_RS <- tbl_df(read.csv2("./data/Export_RS.csv"))
head(DF_RS)[,1:4]
```

## What does the station measure?
```{r EXPcheck, echo=TRUE, warning=FALSE, error=FALSE, comment=""}
levels(DF_RS$component)
```

The variable "component" refers to the measured indicators (NO, NO2, O3, PM10 en PM2.5). 

## Setup the dataset
```{r dataprep, echo=TRUE, warning=FALSE, error=FALSE}
DF_fin <- DF_RS %>%                               ## Create DF-fin
        mutate(comp = as.factor(component)) %>%   ## as factor
        mutate(dt = as.POSIXct(strptime(tijdstip,
                "%Y-%m-%d %H:%M:%S"))) %>%        ## add dt column
        mutate(ym = as.factor(paste(              ## add yr-mon col.
                format(dt, "%Y"), 
                format(dt, "%m"),
                sep="-"))) %>%    
        group_by(ym, comp) %>%                    ## group for mean
        summarise(avg = mean(waarde)) %>%         ## average value
        mutate(avg = round(avg, 2)) %>%           ## round 2 dec's
        group_by() %>%                            ## release groups
        arrange(ym) %>%                           ## sort on yr-mon
        select(ym, comp, avg)                     ## relevant cols
```

## The final dataset
```{r finalset, echo=TRUE, comment=""}
DF_fin
```

## The final plot
```{r leaflet, eval=FALSE, echo=TRUE}
plot_ly(DF_fin, x = ~ym, y = ~avg, color = ~comp, 
        type = 'scatter', mode = 'lines+markers',
        connectgaps=TRUE)
```

## Air Polution (The Hague)
Created: G. Versteeg (July 24th, 2017)

```{r l2, echo=FALSE, warning=FALSE, error=FALSE}
plot_ly(DF_fin, x = ~ym, y = ~avg, color = ~comp, 
        type = 'scatter', mode = 'lines+markers',
        connectgaps=TRUE)
```
